### 2020-03-12 ms/ms.tex
Add Ross's edits

Declines in wild Atlantic salmon (\emph{Salmo salar}) throughout the north
Atlantic are primarily attributed to declining survival at sea. This
hypothesis has proven challenging to test on a river-by-river basis because of
the need to model data on both migrating smolts and returning adults and to
simultaneously estimate multiple parameters, especially for salmon spending
more than one winter at sea (1SW) before spawning. We fit a hierarchical Bayesian
maturity schedule model to data (19-42 years) for seven populations in  
Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
Scotia (NS), Canada. We estimate survival in the first (\So) and second year at sea (\St),
and the proportion returning as 1SW adults (\Pg). Trends in  were not consistent
among rivers. Since 1990, \So increased at Western Arm Brook (NL)
and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivière
de la Trinité (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
declined at LaHave River (NS), and fluctuated at Nashwaak
River (NB). Estimates of \St were highly uncertain, particularly for
1SW-dominated populations; \Pg was generally stable. These results challenge the
narrative that marine survival has changed in a temporally consistent manner
among spatially disparate populations. Our findings suggest that, at the
population level, changes in abundance are attributable to temporal shifts in
multiple components of individual fitness, including at-sea survival. If
salmon populations do not respond in a consistent, uniform manner to changing
environmental conditions throughout their range, future research initiatives
should explore why.

@@

Reductions in fishing mortality, albeit necessary, are not always sufficient
to facilitate population recovery. Experience with numerous commercially
exploited marine fisheries since the early 1990s has shown that not all
populations respond as favourably as anticipated to major reductions in
exploitation \citep{Hutchings2017}. Gradual efforts to close commercial
Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
full moratoria in all regions, beginning in the Maritime provinces (1984) and
following in Newfoundland (1992), Labrador (1998), and Qu\'{e}bec (2000). Since
these closures, many populations have not increased as 
expected \citep{Dempson2004, ICES2019}; some 
have been assessed as considered threatened or endangered by the 
Committee on the Status of Endangered Wildlife in Canada \citep[][]{Cosewic2010}, 
Canada's national science advisory body (to the national government) on
species risk of extinction.
While it is not fully understood what is driving population declines, the potential
drivers of these are many \citep[see ][for a detailed discussion of possible
causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
damming of waterways and changes in the freshwater habitat \citep{Dunfield1985}, acidification
\citep[particularly in the Southern Uplands region of
NS, see][]{Gibson2010}, predation by seals and birds \citep{Cairns2000}, negative
effects of interbreeding or interactions with escaped farmed salmon
\citep{Keyser2018}, and climate-driven changes in survival and productivity \citep{Mills2013}.

@@

Smolt and adult return abundance estimates originate from a variety of
sources. Smolt estimates from the Trinit\'{e}, Saint-Jean, and Conne populations were obtained using a
mark-recapture approach, while estimates from the ..., WAB, and Campbellton populations
were obtained by direct counts using fish counting fences.
For further details on the data collection methodologies refer to 
\citet{Dempson1991, Venoitt2018} for NL populations, 
\citet{April2018}  for QC populations,
\citet{Jones2014} for NB populations,
and \citet{Gibson2009} for NS populations.

Yearly return data are often recorded in terms of two size groups: small ($< 63$ cm
FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
life-history strategies (i.e. 1SW and 2SW), but can be confounded with repeat
spawners of different sizes. To correct for this in returns 
reported as small and large salmon, we estimated the abundance
of 1SW and 2SW returns using yearly scale age data of a subsample of returns:

@@

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate yearly
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome, at least partially, by
using informative priors for all three marine survival parameters in a
Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg).

@@

Trends among populations also varied: Campbellton,
Saint-Jean, and Western Arm Brook populations showed increases in \So
over time, Trinit\'{e}, Conne, and La Have populations showed decreases,
while at Nashwaak there was an increase in median \So during the early
2000s but a decrease in the 2010s. Yearly estimates of \So had very little
variability for one sea-winter dominated populations (Conne, Campbellton, and
WAB), but were more variable (i.e. wider credible intervals) in the other
populations.

@@

% 1. Main findings
Our results challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
Temporal trends are not consistent among populations. 
Over the time periods for which data were available, some rivers show positive trends in survival in the
first winter at sea (\So) while other exhibit highly variable yet stable trends, and some show
declines. We could not assess trends in the second winter at sea (\St) or
proportion returning as grilse (\Pg), as these parameter estimates were highly
uncertain and were strongly influenced by the priors.
Perhaps there is a need to rethink our understanding of Atlantic salmon
population dynamics in light of the possibilities that (1) any real and persistent decline in marine survival was
experienced by some but not necessarily all populations, (2) reductions in survival might
have occurred over a relatively brief period of time and have not persisted, and (3) marine survival has
been relatively stable, or increasing in some populations for one or more decades.

% 2. How they compare to other publications
Our results are contrary to those of \citet{Olmos2019}, who detected positive
correlations in post-smolt survival among spatially broad stock units. These differences
could be due to a number of reasons: different model specifications and
structure, different methods for estimating covariance, difference in the
spatial scales of data sources (i.e. river vs province scales), and perhaps most importantly the use stock-recruitment relationships
rather than empirical smolt count data to estimate marine survival.
As trends in marine survival during the first winter at sea are highly independent
among rivers on relatively small spatial scale, trends from broader
geographical areas (i.e. province, state, or country-wide estimates) may not
be representative.
Interestingly, our estimates of \So and \St are very similar to those produced
by \citet{Chaput2003b}, and our trends are almost identical for the
overlapping time period that marine survival was estimated for in their study
(1984-1998). 
While \citet{Chaput2003b} separated abundance data for males and females
and assumed their survival rates were the same (to be able to reach an
analytical solution), our study reached almost the same results (albeit with
slightly higher uncertainty), using a Bayesian approach with informative
priors. These overlapping trends obtained with two different methods 
suggest that our method is effective at estimating marine survival.

Trends in marine survival
among populations were compared by \citet{Chaput2012a} using adult return rates.
He found that for 4 of 6 populations examined, return rates in the 1990s 
were lower than those during the 1970s.
\citet{Friedland1993} compared return rates for a number of rivers in eastern
North America between 1973 and 1988, and suggested there are similar trends among these. 
However, the similarity in these trends was driven primarily by two years, 1977 and 1978, which
show concurrent low and high relative return rates across rivers,
respectively. Other years are much more variable relative to each other.
\citeauthor{Friedland1993}'s \citeyear{Friedland1993} time series ends in  
1988; thus there are only a few years for which to assess overlap with the
time series in our study.
In any event, we caution that the pooling of adult return rates \citep{Chaput2012a, Friedland1993} 
can mask inter-annual variability in marine survival,
and hence might not produce an accurate depiction of marine survival trends.
\citet{Dempson2003} described a general declining trend in marine survival for
Newfoundland rivers (except WAB); we drew the same conclusion for 
Conne River but not Campbellton River or WAB. It is not possible to draw broader conclusions
with data from only three Newfoundland rivers, but it seems that among index rivers,
those in Newfoundland are among those with the highest marine survival rates.

% 3. Potential reasons
There are a variety of potential explanations for the lack of synchronous
trends in estimates of \So. 
Marine survival in the first winter at sea could be highly variable between
populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation) 
relative to broader-scale, even ocean-wide, drivers.
The synchrony reported for marine survival trends at broader spatial scales \citep{Olmos2019}
might be attributable to the use of stock-recruitment relationships to estimate survival,
relationships that may have been confounded by changes in recruitment dynamics.
There is some evidence of a correlation between return rate and growth (as
indicated by inter-circuli spacing on scales), where years of poor growth
tended to also be years of poor survival \citep{Friedland1993}, supporting the
idea that environmental variability can affect marine survival.
Furthermore, among European salmon, there is evidence of a positive correlation
between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
conditions favour post-smolts \citep{Friedland1998}, based on mapping the
extent of area of suitable temperature (7-13 \textdegree C).

@@

Oceanic conditions have been correlated with abundance trends and growth
\citep{Todd2008}, however, the mechanism by which such bottom-up effects, \
mediated by changes in food availability,
affect population dynamics beyond marine survival needs to
be thoroughly reassessed. If marine survival on its own cannot fully explain
trends in abundance, then there are potential carry-on effects of oceanic
conditions that manifest with regards to fresh production. 
For example, adults
that return to spawn after spending suboptimal conditions at sea might be less
likely to make it to their spawning grounds, successfully secure a mate,
produce fewer eggs, or produce eggs with lower per capita fitness than those
produced by adults which grew in optimal oceanic conditions.
As larger females tend to be more productive, in terms of fecundity and total
reproductive energy, than the same weight's worth of smaller females
\citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
impacts on food availability could potentially have disproportionate effects on fecundity
and fitness of the offspring.

Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}
and changes in the oceanic conditions that spawners experience could be
contributing to this variability.
Obviously, there would be a time lag (perhaps as much as a generation) in how such effects
might be manifest at the adult stage.
However, given that most correlations are between relatively monotonic declines
in abundance coupled also monotonic increases in climatic indices
over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008,
    Beaugrand2012}, it would be expected that this correlations would be
maintained even if salmon abundances were lagged by a generation length.

@@

Secondly, the assumed hierarchical structure might not be the most appropriate
for modelling smolt abundances. This approach results in shrinkage to the
mean, which means that the variability of yearly smolt estimates is less than
it would be if not modelled hierarchically. 
That said, this assumption is likely to result in more conservative trends in marine survival, as the
variation is smolt estimates is reduced.
It is important to note that the hierarchical structure in the estimation of
\Pg seems like a reasonable assumption given that the probability of returning
as grilse has a genetic component associated to it \citep{Aykanat2019} and is
not expected to vary much, within a population, among years.

>>>>

Declines in wild Atlantic salmon (\emph{Salmo salar}) abundance throughout the north
Atlantic are primarily attributed to declining survival at sea. This
hypothesis has proven challenging to test on a river-by-river basis because of
the need to model data on both migrating smolts and returning adults and to
simultaneously estimate multiple parameters, especially for salmon spending
more than one winter at sea (1SW) before spawning. We fit a hierarchical Bayesian
maturity schedule model to data (19-42 years) for seven populations in  
Newfoundland and Labrador (NL), Qu\'{e}bec (QC), New Brunswick (NB), and Nova
Scotia (NS), Canada. We estimate survival in the first (\So) and second year at sea (\St),
and the proportion returning as 1SW adults (\Pg). Trends were not consistent
among rivers. Since 1990, \So increased at Western Arm Brook (NL)
and Rivi\`{e}re Saint-Jean (QC), but declined at Conne River (NL) and Rivi\'{e}re
de la Trinit\'{e} (QC). Since the mid-1990s, \So increased at Campbellton River (NL),
declined at LaHave River (NS), and fluctuated at Nashwaak
River (NB). Estimates of \St were highly uncertain, particularly for
1SW-dominated populations; \Pg was generally stable. These results challenge the
narrative that marine survival has changed in a temporally consistent manner
among spatially disparate populations. Our findings suggest that, at the
population level, changes in abundance are attributable to temporal shifts in
multiple components of individual fitness, including at-sea survival. If
salmon populations do not respond in a consistent, uniform manner to changing
environmental conditions throughout their range, future research initiatives
should explore why.

@@

Reductions in fishing mortality, albeit necessary, are not always sufficient
to facilitate population recovery. Experience with numerous commercially
exploited marine fisheries since the early 1990s has shown that not all
populations respond as favourably as anticipated to major reductions in
exploitation \citep{Hutchings2017}. Gradual efforts to close commercial
Atlantic salmon (\emph{Salmo salar}) fisheries in eastern Canada culminated in
full moratoria in all regions, beginning in the Maritime provinces (1984) and
following in Newfoundland (1992), Labrador (1998), and Qu\'{e}bec (2000). Since
these closures, many populations have not increased as 
expected \citep{Dempson2004, ICES2019}; some 
have been assessed as threatened or endangered by the 
Committee on the Status of Endangered Wildlife in Canada \citep[][]{Cosewic2010}, 
Canada's science advisory body (to the national government) on
species risk of extinction.
While it is not fully understood what is driving population declines, the potential
drivers of these are many \citep[see ][for a detailed discussion of possible
causes]{Cairns2001}, including but not limited to: fishing mortality \citep{Dempson2004}, 
damming of waterways and changes in the freshwater habitat \citep{Dunfield1985,Clarke2014a}, acidification
\citep[particularly in the Southern Uplands region of
NS, see][]{Gibson2010}, predation by seals and birds \citep{Cairns2000}, negative
effects of interbreeding or interactions with escaped farmed salmon
\citep{Keyser2018}, and climate-driven changes in survival and productivity \citep{Mills2013}.

@@

Smolt and adult return abundance estimates originate from a variety of
sources. Smolt estimates from the Trinit\'{e}, Saint-Jean, and Conne populations were obtained using a
mark-recapture approach, while estimates from the ..., WAB, and Campbellton populations
were obtained by direct counts using fish counting fences.
For further details on the data collection methodologies refer to 
\citet{Dempson1991, Venoitt2018} for NL populations, 
\citet{April2018}  for QC populations,
\citet{Jones2014} for NB population,
and \citet{Gibson2009} for NS population.

Annual return data are often recorded in terms of two size groups: small ($< 63$ cm
FL) and large ($\geq 63$ cm FL) salmon, as these closely represent different
life-history strategies (i.e. 1SW and 2SW), but can be confounded with repeat
spawners of different sizes. To correct for this in returns 
reported as small and large salmon, rather than 1SW and 2SW fish, we estimated the abundance
of 1SW and 2SW returns using yearly scale age data of a subsample of returns:

@@

We developed a hierarchical Bayesian model that uses Murphy's maturity
schedule method, in conjunction with informative priors, to estimate annual
marine survival in seven populations of Atlantic salmon. We account for
observation error in smolt and return estimates, as well as estimating the
proportion returning after one winter (i.e. \Pg) hierarchically.
There is an identifiability problem in the maturity schedule equations where
the parameter estimates cannot be optimally solved \citep{Chaput2003a}.
However, this issue can be mathematically overcome, at least partially, by
using informative priors for all three marine survival parameters in a
Bayesian framework.
This method requires abundance estimates of smolts as well as abundance estimates
of returning one-sea-winter (1SW) and two-sea-winter adults. With these data,
it estimates three parameters: survival in the first year at sea (\So), survival
in the second year at sea (\St), and the proportion of fish returning after one
year at sea (\Pg).

@@

Trends among populations also varied: Campbellton,
Saint-Jean, and Western Arm Brook populations showed increases in \So
over time, Trinit\'{e}, Conne, and LaHave populations showed decreases,
while at Nashwaak there was an increase in median \So during the early
2000s but a decrease in the 2010s. Yearly estimates of \So had very little
variability for one sea-winter dominated populations (Conne, Campbellton, and
WAB), but were more variable (i.e. wider credible intervals) in the other
populations.

@@

% 1. Main findings
Our results challenge the narrative that marine survival, specifically survival in the first year at sea, is declining
uniformly throughout the range of Atlantic salmon in the northwest Atlantic.
Temporal trends are not consistent among all populations. 
Over the time periods for which data were available, some rivers show positive trends in survival in the
first winter at sea (\So) while other exhibit highly variable yet stable trends, and some show
declines. We could not assess trends in the second winter at sea (\St) or
proportion returning as grilse (\Pg), as these parameter estimates were highly
uncertain and were strongly influenced by the priors.
Perhaps there is a need to rethink our understanding of Atlantic salmon
population dynamics in light of the possibilities that (1) any real and persistent decline in marine survival was
experienced by some but not necessarily all populations, (2) reductions in survival might
have occurred over a relatively brief period of time and have not persisted, and (3) marine survival has
been relatively stable, or increasing in some populations for one or more decades.

% 2. How they compare to other publications
Our results are contrary to those of \citet{Olmos2019}, who detected mostly positive
correlations in post-smolt survival among spatially broad stock units. These differences
could be due to a number of reasons: different model specifications and
structure, different methods for estimating covariance, difference in the
spatial scales of data sources (i.e. river vs province scales), and perhaps most importantly the use stock-recruitment relationships
rather than empirical smolt count estimates to estimate marine survival.
As trends in marine survival during the first winter at sea are highly independent
among rivers on relatively small spatial scale, trends from broader
geographical areas (i.e. province, state, or country-wide estimates) may not
be representative.
Interestingly, our estimates of \So and \St are very similar to those produced
by \citet{Chaput2003b}, and our trends are almost identical for the
overlapping time period that marine survival was estimated for in their study
(1984-1998). 
While \citet{Chaput2003b} separated abundance data for males and females
and assumed their survival rates were the same (to be able to reach an
analytical solution), our study reached almost the same results (albeit with
slightly higher uncertainty), using a Bayesian approach with informative
priors. These overlapping trends obtained with two different methods 
suggest that our method is effective at estimating marine survival.

Trends in marine survival
among populations were compared by \citet{Chaput2012a} using adult return rates.
He found that for 4 of 6 populations examined, return rates in the 1990s 
were lower than those during the 1970s.
\citet{Gibson2016} calculated higher return rates of 2SW for the Nashwaak in
the 1970s than in the period since, with return rates of 1SW in the 1970s
being comparable to those in the late 2000s.
\citet{Friedland1993} compared return rates for a number of rivers in eastern
North America between 1973 and 1988, and suggested there are similar trends among these. 
However, the similarity in these trends was driven primarily by two years, 1977 and 1978, which
show concurrent low and high relative return rates across rivers,
respectively. Other years are much more variable relative to each other.
\citeauthor{Friedland1993}'s \citeyear{Friedland1993} time series ends in  
1988; thus there are only a few years for which to assess overlap with the
time series in our study.
While declines in return rates since the 1970s seem to be consistent among populations, 
we were unable to assess if marine survival estimates were also higher in the 1970s as we lacked 
data from this decade.
In any event, we caution that the pooling of adult return rates \citep{Chaput2012a, Friedland1993,Gibson2016} 
can mask inter-annual variability in marine survival,
and hence might not produce an accurate depiction of marine survival trends.
\citet{Dempson2003} described a general declining trend in marine survival for
Newfoundland rivers (except WAB); we drew the same conclusion for 
Conne River but not Campbellton River or WAB. It is not possible to draw broader conclusions
with data from only three Newfoundland rivers, but it seems that among index rivers,
those in Newfoundland are among those with the highest marine survival rates.

% 3. Potential reasons
There are a variety of potential explanations for the lack of synchronous
trends in estimates of \So. 
Marine survival in the first winter at sea could be highly variable between
populations because of the predominance of spatially local environmental drivers of survival (e.g., temperature, predation, prey abundance) 
relative to broader-scale, even ocean-wide, drivers.
The synchrony reported for marine survival trends at broader spatial scales \citep{Olmos2019}
might be attributable to the use of stock-recruitment relationships to estimate survival,
relationships that may have been confounded by changes in recruitment dynamics.
There is some evidence of a correlation between return rate and growth (as
indicated by inter-circuli spacing on scales), where years of poor growth
tended to also be years of poor survival \citep{Friedland1993}, supporting the
idea that environmental variability can affect marine survival.
Furthermore, among European salmon, there is evidence of a positive correlation
between spring temperature in the Norwegian and North Seas and population abundance, suggesting warmer
conditions favour post-smolts \citep{Friedland1998}, based on mapping the
extent of area of suitable temperature (7-13 \textdegree C).

@@

Oceanic conditions have been correlated with abundance trends and growth
\citep{Todd2008}, however, the mechanism by which such bottom-up effects, \
mediated by changes in food availability,
affect population dynamics beyond marine survival needs to
be thoroughly reassessed. If marine survival on its own cannot fully explain
trends in abundance, then there are potential carry-on effects of oceanic
conditions that manifest with regards to freshwater production. 
For example, adults
that return to spawn after spending suboptimal conditions at sea might be less
likely to make it to their spawning grounds, successfully secure a mate,
produce fewer eggs, or produce eggs with lower per capita fitness than those
produced by adults which grew in optimal oceanic conditions.
As larger females tend to be more productive, in terms of fecundity and total
reproductive energy, than the same weight's worth of smaller females
\citep{Barneche2018}, a small decrease in body condition resulting from bottom-up
impacts on food availability and quality could potentially have disproportionate effects on fecundity
and fitness of the offspring.

Egg-to-smolt survival in Atlantic salmon is highly variable \citep{Klemetsen2003,Chaput2015}.

While this variability is likely attributable to changes in freshwater
conditions (e.g. discharge, temperature, water quality) and uncertainty in
spawner and smolt estimates, 
changes in the oceanic conditions that spawners experience could also be
contributing to this variability.
Obviously, there would be a time lag (perhaps as much as a generation) in how such effects
might be manifest at the adult stage.
However, given that most correlations are between relatively monotonic declines
in abundance coupled also monotonic increases in climatic indices
over decadal time scales \citep[e.g.,][]{Friedland1998, Todd2008,
    Beaugrand2012}, it would be expected that these correlations would be
maintained even if salmon abundances were lagged by a generation length.

@@

Secondly, the assumed hierarchical structure might not be the most appropriate
for modelling smolt abundances. This approach results in shrinkage to the
mean, which means that the variability of annual smolt estimates is less than
it would be if not modelled hierarchically. 
That said, this assumption is likely to result in more conservative trends in marine survival, as the
variation is smolt estimates is reduced.
It is important to note that the hierarchical structure in the estimation of
\Pg seems like a reasonable assumption given that the probability of returning
as grilse has a genetic component associated to it \citep{Aykanat2019} and is
not expected to vary much, within a population, among years.
